---
title: "MGSC Final Project"
author: "Anjali Maniar and Alyssa Ruste"
subtitle: MGSC 310 
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

sessionInfo()
getwd()

```



```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
library('ggplot2')
library('fastDummies')
library('AMR')
library('dplyr')
library('partykit')
library('PerformanceAnalytics')
library('rpart')       
library('rpart.plot')  
library('randomForest')
library('randomForestExplainer')
library('caret')
library('e1071')
library('Metrics')
library('glmnet')
library('glmnetUtils')

```


```{r}

airplane_train <- read.csv(here::here("datasets", "plane_train.csv"))

airplane_train$Sex <-
  ifelse(airplane_train$Gender == "Female", 1, 0)

airplane_train$Types.of.Customers <-
  ifelse(airplane_train$Customer.Type == "Loyal Customer", 1, 0)

airplane_train$Travel.Types <-
  ifelse(airplane_train$Type.of.Travel == "Personal Travel", 1, 0)

airplane_train$satisfaction <-
  ifelse(airplane_train$satisfaction == "satisfied", 1, 0)

airplane_train <- dummy_cols(airplane_train, select_columns = 'Class')

airplane_train$age_grouping <- cut(airplane_train$Age, breaks = c(0, 18, 29, 39, 49, 64, 100),
labels = c("0-18yrs", "19-29yrs", "30-39yrs", "40-49yrs","50-64yrs", "65+ yrs"))

airplane_train$satisfied <- factor(airplane_train$satisfaction, labels=c("No", "Yes"))

airplane_train
```

```{r}
airplane_train <-
  airplane_train %>%
  mutate(Class = as.factor(Class))

airplane_train$Class <- relevel(airplane_train$Class, ref = "Eco")


airplane_train <-
  airplane_train %>%
  mutate(Inflight.wifi.service = as.factor(Inflight.wifi.service),
         Departure.Arrival.time.convenient = as.factor(Departure.Arrival.time.convenient),
         Ease.of.Online.booking = as.factor(Ease.of.Online.booking),
         Gate.location = as.factor(Gate.location),
         Food.and.drink = as.factor(Food.and.drink),
         Online.boarding = as.factor(Online.boarding),
         Seat.comfort = as.factor(Seat.comfort),
         Inflight.entertainment = as.factor(Inflight.entertainment),
         On.board.service = as.factor(On.board.service),
         Leg.room.service = as.factor(Leg.room.service),
         Baggage.handling = as.factor(Baggage.handling),
         Checkin.service = as.factor(Checkin.service),
         Inflight.service = as.factor(Inflight.service),
         Cleanliness = as.factor(Cleanliness),
         age_grouping = as.factor(age_grouping),
         Gender = as.factor(Gender))

head(airplane_train)

```

```{r}
airplane_test <- read.csv(here::here("datasets", "plane_test.csv"))

airplane_test$Sex <-
  ifelse(airplane_test$Gender == "Female", 1, 0)

airplane_test$Types.of.Customers <-
  ifelse(airplane_test$Customer.Type == "Loyal Customer", 1, 0)

airplane_test$Travel.Types <-
  ifelse(airplane_test$Type.of.Travel == "Personal Travel", 1, 0)

airplane_test$satisfaction <-
  ifelse(airplane_test$satisfaction == "satisfied", 1, 0)

airplane_test <- dummy_cols(airplane_test, select_columns = 'Class')

airplane_test$age_grouping <- cut(airplane_test$Age, breaks = c(0, 18, 29, 39, 49, 64, 100),
labels = c("0-18yrs", "19-29yrs", "30-39yrs", "40-49yrs","50-64yrs", "65+ yrs"))

airplane_test$satisfied <- factor(airplane_test$satisfaction, labels=c("No", "Yes"))

airplane_test <-
  airplane_test %>%
  mutate(Class = as.factor(Class))

airplane_test$Class <- relevel(airplane_test$Class, ref = "Eco")


airplane_test <-
  airplane_test %>%
  mutate(Inflight.wifi.service = as.factor(Inflight.wifi.service),
         Departure.Arrival.time.convenient = as.factor(Departure.Arrival.time.convenient),
         Ease.of.Online.booking = as.factor(Ease.of.Online.booking),
         Gate.location = as.factor(Gate.location),
         Food.and.drink = as.factor(Food.and.drink),
         Online.boarding = as.factor(Online.boarding),
         Seat.comfort = as.factor(Seat.comfort),
         Inflight.entertainment = as.factor(Inflight.entertainment),
         On.board.service = as.factor(On.board.service),
         Leg.room.service = as.factor(Leg.room.service),
         Baggage.handling = as.factor(Baggage.handling),
         Checkin.service = as.factor(Checkin.service),
         Inflight.service = as.factor(Inflight.service),
         Cleanliness = as.factor(Cleanliness),
         age_grouping = as.factor(age_grouping),
         Gender = as.factor(Gender))

head(airplane_test)

```

```{r}
levels(airplane_test$Gender) <- levels(airplane_train$Gender)
levels(airplane_test$Class) <- levels(airplane_train$Class)
levels(airplane_test$Inflight.wifi.service) <- levels(airplane_train$Inflight.wifi.service)
levels(airplane_test$Departure.Arrival.time.convenient) <- levels(airplane_train$Departure.Arrival.time.convenient)
levels(airplane_test$Ease.of.Online.booking) <- levels(airplane_train$Ease.of.Online.booking)
levels(airplane_test$Gate.location) <- levels(airplane_train$Gate.location)
levels(airplane_test$Food.and.drink) <- levels(airplane_train$Food.and.drink)
levels(airplane_test$Online.boarding) <- levels(airplane_train$Online.boarding)
levels(airplane_test$Seat.comfort) <- levels(airplane_train$Seat.comfort)
levels(airplane_test$Inflight.entertainment) <- levels(airplane_train$Inflight.entertainment)
levels(airplane_test$On.board.service) <- levels(airplane_train$On.board.service)
levels(airplane_test$Leg.room.service) <- levels(airplane_train$Leg.room.service)
levels(airplane_test$Baggage.handling) <- levels(airplane_train$Baggage.handling)
levels(airplane_test$Checkin.service) <- levels(airplane_train$Checkin.service)
levels(airplane_test$Inflight.service) <- levels(airplane_train$Inflight.service)
levels(airplane_test$Cleanliness) <- levels(airplane_train$Cleanliness)
levels(airplane_test$age_grouping) <- levels(airplane_train$age_grouping)
levels(airplane_test$satisfied) <- levels(airplane_train$satisfied)


```

```{r}
summary(airplane_train)
```

  
Ggplot 1: 
```{r}
ggplot(airplane_train, 
       aes(x = Class, 
           fill = satisfied)) + 
  geom_bar(position = "stack")

```
Ggplot 2:
```{r}
ggplot(airplane_train, aes(x=Type.of.Travel, y=Flight.Distance, color=Gender)) + geom_boxplot() 

```

Ggplot 3:
```{r}
ggplot(airplane_train, aes(x=satisfied, y=Age, color=Gender)) + geom_boxplot() 

```


Alyssa: Estimate a random forest model predicting satisfaction as a function of Gender, age_grouping,  Types.of.Customers, Travel.Types, Class, Inflight.wifi.service, Departure.Arrival.time.convenient, Ease.of.Online.booking, Gate.location, Food.and.drink, Online.boarding, Seat.comfort, Inflight.entertainment, On.board.service, Leg.room.service, Baggage.handling, Checkin.service, Inflight.service, Cleanliness, Departure.Delay.in.Minutes and Arrival.Delay.in.Minutes.
```{r}
airplane_sliced <- 
  airplane_train %>% slice(1:8000)

airplane_mod <- randomForest(satisfaction ~ Gender + age_grouping +  Types.of.Customers + Travel.Types + Class +Inflight.wifi.service + Departure.Arrival.time.convenient + Ease.of.Online.booking + Gate.location + Food.and.drink + Online.boarding + Seat.comfort + Inflight.entertainment + On.board.service + Leg.room.service + Baggage.handling + Checkin.service + Inflight.service + Cleanliness + Departure.Delay.in.Minutes + Arrival.Delay.in.Minutes,
                             data = airplane_sliced,
                       type = classification,
                       mtry = 3,
                       na.action = na.roughfix,
                       ntree = 600, 
                       importance = TRUE)

```

Use the plot function on the fitted object to produce the tree plot 
```{r}

plot(airplane_mod)
```

Tune to find the best mtry.
```{r}
tuneGrid <- expand.grid(.mtry = c(1: 10))
rf_mtry <- randomForest(satisfaction ~ Gender + age_grouping + Sex +  Types.of.Customers + Travel.Types + Class +Inflight.wifi.service + Departure.Arrival.time.convenient + Ease.of.Online.booking + Gate.location + Food.and.drink + Online.boarding + Seat.comfort + Inflight.entertainment + On.board.service + Leg.room.service + Baggage.handling + Checkin.service + Inflight.service + Cleanliness + Departure.Delay.in.Minutes + Arrival.Delay.in.Minutes,
                             data = airplane_sliced,
    method = "rf",
    metric = "Accuracy",
    tuneGrid = tuneGrid,
    trControl = trControl,
    importance = TRUE,
    nodesize = 14,
    ntree = 200,
    na.action=na.roughfix)
print(rf_mtry)

```

Use new mtry and ntrees best fit for model.
```{r}

airplane_sliced <- 
  airplane_train %>% slice(1:8000)

airplane_modnew <- randomForest(satisfaction ~ Gender + age_grouping +  Types.of.Customers + Travel.Types + Class +Inflight.wifi.service + Departure.Arrival.time.convenient + Ease.of.Online.booking + Gate.location + Food.and.drink + Online.boarding + Seat.comfort + Inflight.entertainment + On.board.service + Leg.room.service + Baggage.handling + Checkin.service + Inflight.service + Cleanliness + Departure.Delay.in.Minutes + Arrival.Delay.in.Minutes,
                             data = airplane_sliced,
                       type = classification,
                       mtry = 7,
                       na.action = na.roughfix,
                       ntree = 200, 
                       importance = TRUE)
```

Use the print function against the fitted model to view the text descriptions of the model fit
```{r}
print(airplane_modnew)

```

Plot of multiimportance frame for variables.
```{r}
importance_frame <- measure_importance(airplane_modnew)
plot_multi_way_importance(importance_frame, size_measure = "no_of_nodes")

```

Testing performance of model on train set. 
```{r}
prediction<-predict(airplane_modnew,airplane_train) 
postResample(pred = prediction, obs = airplane_train$satisfaction)
```

Testing performance of model on test set.
```{r}
prediction<-predict(airplane_modnew,airplane_test) 
postResample(pred = prediction, obs = airplane_test$satisfaction)
```

Variable importance
```{r}
varImp(airplane_modnew)
```

Lasso Logistic Regression(Anjali):
```{r}

lasso_mod <- cv.glmnet(satisfaction ~ Gender + age_grouping +  Types.of.Customers + Travel.Types + Class +Inflight.wifi.service + Departure.Arrival.time.convenient + Ease.of.Online.booking + Gate.location + Food.and.drink + Online.boarding + Seat.comfort + Inflight.entertainment + On.board.service + Leg.room.service + Baggage.handling + Checkin.service + Inflight.service + Cleanliness + Departure.Delay.in.Minutes + Arrival.Delay.in.Minutes, data = airplane_train, alpha = 1)



lasso_coefs <- data.frame(
  `lasso_min` = coef(lasso_mod, s = lasso_mod$lambda.min) %>% 
    round(3) %>% as.matrix() %>% as.data.frame(),
  `lasso_1se` = coef(lasso_mod, s = lasso_mod$lambda.1se) %>% 
    round(3) %>% as.matrix() %>%  as.data.frame())  %>% rename(`lasso_min` = 1, `lasso_1se` = 2)

#coef(lasso_mod)
print(lasso_coefs)

```

```{r}
print(lasso_mod$lambda.min)
print(lasso_mod$lambda.1se)
print(exp(lasso_mod$lambda.min))
print(exp(lasso_mod$lambda.1se))

``` 

```{r}
plot(lasso_mod)

``` 

```{r}

coef(lasso_mod,  s= "lambda.min")

```  

MSE Test:
```{r}

airplane_test_complete <- na.omit(airplane_test)

y_pred <- predict(lasso_mod, newdata = airplane_test)
library('Metrics')
mse(y_pred, airplane_test_complete$satisfaction)
``` 

MSE Train:
```{r}
airplane_train_complete <- na.omit(airplane_train)
y_pred_train <- predict(lasso_mod, newdata = airplane_train_complete)
mse(y_pred_train, airplane_train_complete$satisfaction)


``` 

MAE Test:
```{r}

airplane_test_complete <- na.omit(airplane_test)

y_pred <- predict(lasso_mod, newdata = airplane_test)
library('Metrics')
mae(y_pred, airplane_test_complete$satisfaction)
``` 

MAE Train:
```{r}
airplane_train_complete <- na.omit(airplane_train)
y_pred_train <- predict(lasso_mod, newdata = airplane_train_complete)
mae(y_pred_train, airplane_train_complete$satisfaction)


``` 

Coefficients:
```{r}
lasso_coefs <- data.frame(
  `lasso_min` = coef(lasso_mod, s = lasso_mod$lambda.min) %>% 
    round(3) %>% as.matrix() %>% as.data.frame(),
  `lasso_1se` = coef(lasso_mod, s = lasso_mod$lambda.1se) %>% 
    round(3) %>% as.matrix() %>%  as.data.frame())  %>% rename(`lasso_min` = 1, `lasso_1se` = 2)


print(lasso_coefs)
plot(lasso_mod)

print(lasso_mod)

``` 


